Privacy-Respecting Type Error Telemetry at Scale

### 2023-12-12 programming23/paper.tex
ctc: developer not necessarily to blame
revisions

\newcommand{\etal}{et~al.}
\newcommand{\uid}{\textsc{id}}
\newcommand{\IDE}{\textsc{ide}}
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\FILL}{\textbf{FILL}}
\newcommand{\dotscale}[1]{\scalebox{0.72}{#1}}
\newcommand{\wideas}[2]{\makebox[\widthof{#2}][l]{#1}}
\newcommand{\twoline}[2]{\parbox[s]{1.4cm}{\flushleft#1\newline#2}}
\newcommand{\chkYes}{\dotscale{\CIRCLE}}
\newcommand{\chkMaybe}{\wideas{\dotscale{\Circle}}{\chkYes}}
\newcommand{\chkNo}{\wideas{}{\chkYes}}
\newcommand{\pct}[1]{\SI{#1}{\percent}}
\newcommand{\percentile}[1]{{#1}th~percentile}
\newcommand{\ptile}[1]{\textsc{p}{#1}}
\newcommand{\modefont}[1]{\texttt{#1}}
\newcommand{\mnocheck}{\modefont{nocheck}}
\newcommand{\mnonstrict}{\modefont{nonstrict}}
\newcommand{\mstrict}{\modefont{strict}}
\newcommand{\FS}{background}
\newcommand{\zerowidth}[1]{\makebox[0pt][l]{#1}}
\newcommand{\zeroheight}[1]{\raisebox{0pt}[0pt][0pt]{\(#1\)}}
\newcommand{\nK}[1]{#1\textsc{k}}
\newcommand{\stddev}[1]{\nK{#1}}
\newcommand{\gcell}[1]{\cellcolor{green!20}#1}
\newcommand{\ycell}[1]{\cellcolor{yellow!18}#1}
\newcommand{\ocell}[1]{\cellcolor{orange!29}#1}
\newcommand{\rcell}[1]{\cellcolor{red!30}$\!\!$#1$\!\!$}
\newcommand{\gbox}[1]{\colorbox{green!20}{#1}}
\newcommand{\ybox}[1]{\colorbox{yellow!18}{#1}}
\newcommand{\obox}[1]{\colorbox{orange!29}{#1}}
\newcommand{\rbox}[1]{\colorbox{red!30}{#1}}
\newcommand{\tefsfont}[1]{\textsc{#1}}
\newcommand{\tekey}{\tefsfont{te}}
\newcommand{\fskey}{\tefsfont{bg}}
\newcommand{\QALAN}{\FILL{} Alan:}
\newcommand{\panon}{pseudonymized}
\newcommand{\symbox}[1]{{\makebox[0.8em][c]{{#1}}}}
\newcommand{\addsym}{\symbox{\(\uparrow\)}}
\newcommand{\keepsym}{\symbox{\(=\)}}
\newcommand{\dropsym}{\symbox{\(\downarrow\)}}
% \newcommand{\dsymbox}[1]{\makebox[0.8em][c]{#1}}
% \newcommand{\dropsym}{\dsymbox{\raisebox{0ex}[0.9em]{\raisebox{0.5ex}{\vrule width 0.6em height 0.6pt}}}}

@@

Due to the broad community of creators, the goals of the Luau type
system are rather unique~\cite{bfj-hatra-2021}.
Whereas a traditional type system focuses on compilation and memory
safety, Luau takes a lenient approach by default and lets creators
gradually~\cite{st-sfp-2006,tf-popl-2008} migrate to rigorous checks
one module at a time.
In this way, for example,
students in code camps may begin in the \mnocheck{} mode that reports only
syntax errors; hobbyists can use \mnonstrict{} mode to identify
would-be runtime errors; and developers who need to ensure software
quality can rely on \mstrict{} mode.
(No matter the mode, a {\FS{}} type analysis still drives the \IDE{} tools.)
Luau's types aim to support untyped designs, in the spirit of migratory
typing~\cite{tfffgksst-snapl-2017}, so that creators can switch modes without
needing to restructure their code in major, potentially-breaking, ways.

@@

  \item
    An available dataset of over 1.5 million telemetry records
    and scripts to analyze them.
    \emph{TODO We intend to submit these in an artifact.}

@@

  \begin{tabular}{ll}
    Label & Interpretation \\\midrule
    \code{TypeMismatch} & Basic type error. \\
    \code{SyntaxError} & Basic parse error, e.g., \code{for if end}. \\
    \code{UnknownProperty} & Referenced an invalid field or method.  \\
    \code{OnlyTablesCanHaveMethods} & Tried to attach a method to a non-table. \\
    \code{CannotExtendTable} & Tried to extend a sealed table. \\
    \code{TypesAreUnrelated} & Failed cast, unify, or subtype. \\
    \code{CountMismatch} & Arity mismatch for a function. \\
    %% \code{CannotCallNonFunction} & Called a value that is not a function. \\
    \code{IncorrectGenericParamCount} & Arity mismatch for a generic type. \\
    \code{CodeTooComplex} & Type analysis failed, cannot understand the code. \\
    %% \code{UnificationTooComplex} & Type analysis failed, unification solver hit a limit \\
    \code{GenericError}
    & Generic label for other non-type errors,\\
    & \hbox{}~~such as looping over an unordered table.
    %% attempting to extend a type that does not describe a class
    %% attempting to iterate over a table without a clear order

@@

  \item
    Telemetry \emph{must transmit a small amount of data}---not too many records
    with not too much information inside---to avoid overloading the servers.

\end{enumerate}
Cox~\cite{transparent-telemetry} discusses in detail the privacy implications
and tradeoffs in telemetry.

In addition, due to the architecture of Roblox Studio,
there is some information that is not available to the
typechecker:
\begin{itemize}
  \item
    Lifecycle events including as save, run, quit, and publish.\footnote{Publish events
    cannot be reported at any rate because they are public, and could be used
    to match a telemetry session to a creator.}
  \item
    GUI state: in particular, whether the Script Analysis widget is visible.
\end{itemize}
Our type error telemetry has to meet all of the above constraints.

@@

\Cref{f:tdata} shows the structure of our telemetry records:
\begin{enumerate}
  \item
    \uid{}: a \panon{} identifier (16-digit random number) for the current session.
  \item
    Time: one timestamp from the client and one from the server.
    %% timestamp explanation \cref{s:data-cleaning}
  \item
    Mode: type analysis mode of the current file: \mnocheck{},
    \mnonstrict{}, or \mstrict{}.
  \item
    Reason: a flag that explains whether this record was sent due to a
    randomly-selected keystroke or a module switch.
  \item
    Size Info: number of lines in the codebase and number of lines in the edit range.
  \item
    Overall Counts: summarizes type errors and \FS{} errors during the
    last two invocations of type analysis, and additionally reports the number
    of times type analysis hit an internal limit on problem
    size~(see~\cref{s:code-too-complex} for details).
    For each invocation and each kind of
    analysis error, there are three summary counts:
    %% grand total = 12
    \begin{itemize}
      \item total number of errors,
      \item errors in the current module, and
      \item errors in the current edit range.
    \end{itemize}
  \item
    Edit Range Counts: these list specific errors that arose in the last two
    type analyses and overlap with the current edit range.
    Since there are 35 possible errors~(\cref{s:context}), there can be up to 70 counts
    in a telemetry record.
    Exactly how to interpret this data depends on which analysis it came from:
    \begin{itemize}
      \item
        Errors in the latest type analysis are currently visible to the
        creator.
        These may have been introduced by the changes in the edit range.
      \item
        Errors from the previous type analysis were visible before the creator
        made the latest edits.
        These may have motivated some of the changes in the edit range.
    \end{itemize}
\end{enumerate}

To track the edit range, we record a start and end position, which we
update appropriately on every edit. This can result in very large and
imprecise edit ranges, for example, if the user edits at the beginning and end
of the file.
The upshot of this strategy is that it reduces the size and complexity of
telemetry records because there is only one interval for errors to overlap with.

This telemetry design is admittedly coarse-grained.
For instance, it does not distinguish edits that ignore a type
error from edits that remove one error while introducing another.
We acknowledge this and other threats in~\cref{s:threats}.
The main advantages of this telemetry are the small size of each record
and the complete lack of private information.
Furthermore, despite the limitations, this telemetry supports a variety of
inferences that we showcase in the next section.

@@

\Cref{f:records-per-hour} provides a time-ordered distribution of the data.
The x-axis is labeled with a California time zone; the thin vertical lines
mark the start of a new day.
For each of the 24 hours in each day, there is one blue bar
indicating the number of records generated during that hour.
Each hour has a few dozen to over 3,000 records, typically about
1,000.
{We conjecture the very low bars are due to outage on
  the logging server, but need to investigate further.}
The tall bars are FILL TODO.
%% \FILL{} investigate one tower, what's happening?? all one session?
Telemetry ended in early April; the long tail is due to developers
updating Roblox Studio later.
% BEN: I see no basis for this; nor is this at all interesting.
%% ben: basis = it's where the peaks are
% We do not know which timezone the records originated in, but since the counts
% tend to peak midday California time it seems that the majority
% of Roblox developers are following a Western Hemisphere schedule.
The regular pattern of the peaks suggests that many creators
follow a regular schedule; these may be
professional developers.
The tall weekend (shaded) regions could be because
Roblox has a significant school-aged creator community.

@@

There is a huge amount of variation across projects.
The largest ones have over 50,000 lines of code
while the smallest have zero lines of code.
%% TODO filter the zero lines? what happens in that data?? run a query
Unsurprisingly, these wide-ranging numbers come with large standard
deviations.
The median values are more reasonable, with roughly 3,000 lines of code
and 1,000 lines in edit ranges.

@@

\Cref{t:count-analysis-errors} lists the number of type analysis errors
and breaks them down by location in the codebase.
In total, there are over 590,000 type errors.
The number of \FS{} errors is much larger, at 72 million, because this analysis
uses rigorous checks and runs on every module no matter what type analysis
mode the module declares.
(Furthermore, \FS{} checks run silently, so creators have no
awareness of or incentive to fix these errors.)
Approximately half the analysis errors occurred in the current module,
and a small fraction overlapped with the current edit range~(\pct{5} for
type errors, \pct{1} for \FS{}).

The large fraction of errors in the current module is encouraging
for two reasons.
First, it suggests that errors appear locally, as the result of edits to nearby
code, rather than as the result of edits to code that lives in another
module.
Second, it suggests that creators typically fix errors before switching
to another module.

The low fraction of errors in the edit range sends a mixed message:
\begin{itemize}
  \item
    On one hand, it suggests that edits successfully remove errors.
    This is true, however, only for type errors and not for \FS{} errors.
    There are 11,479 more type errors from the previous analysis
    that overlap with the current edit range~(\pct{27} difference),
    but 172,678 fewer previous \FS{} errors~(\pct{15} difference).
    Since creators can see the type errors but not the \FS{}
    errors before making edits, we conclude that the visibility
    makes a difference.
    \Cref{s:type-error-survival} explores edits and type errors in more detail.

@@

The data does not support this explanation.
On average, mode-upgrades resulted in 3 additional
type errors (stddev 7, median 0).
The worst-case increase was quite high at 57 type errors, but exceptional.
In another exception, upgrading modes removed three type errors,
possibly due to other edits being rolled in with the
mode change.

@@

Digging further into \FS{} errors, the median number of such errors by mode is
1 in \mnocheck{}, 3 in \mnonstrict{}, and 1 in \mstrict{}.
Records in \mnocheck{} mode evidently do not have significantly more \FS{}
errors.
Furthermore, type analysis does not seem to reduce \FS{}
errors; this is surprising and~\cref{s:tefs-compass} has more to say about it.

@@

The main focus of our telemetry design is to learn how type errors intersect
with edit ranges.
Indeed, most of the fields in a telemetry record are dedicated to this
topic~(\cref{f:tdata}), and are directed to RQ2 on type errors
(\cref{s:introduction}).

@@

The counts are imprecise because of telemetry granularity.
If a creator removes one error but then
introduces another on a different line, this typically manifests as
a \keepsym{} rather than one \addsym{} and one \dropsym{}.

@@

  \item
    \code{SyntaxError} and the related \code{UnknownSymbol} and
    \code{UnknownProperty} are the most common.
    If the edit range errors are indicative of type errors at large,
    then most errors in the data are likely due to typos.
    These errors also survive (\keepsym{}) the most often.

    Our telemetry strategy likely inflates
    the number of typos because it cannot determine when a creator is
    mid-edit.
    For example, a creator who writes a method call
    \code{bucket.countFish()} letter-by-letter will generate an
    \code{UnknownSymbol} error for \code{buck} and
    \code{UnknownProperty} for \code{bucket.c}, which may end up sampled.
    \Cref{s:conclusion} suggests ways
    to avoid this in the future.

@@

\begin{itemize}
  \item
    Syntax errors, in particular \code{UnknownSymbol} (for an unbound variable),
    are much more common in \mnonstrict{} than \mstrict{}.
    This may be because \mnonstrict{} reports fewer errors
    to dilute the proportion of unknown symbols.

  \item
    Despite being less strict, \mnonstrict{} reported four errors that
    never appeared in the \mstrict{} records:
    \code{DuplicateTypeDefinition}, \code{OnlyTablesCanHaveMethods},
    \code{MissingUnionProperty}, and \code{FunctionDoesNotTakeSelf}.
    We attribute this to the much larger number of \mnonstrict{} records in the
    dataset.

@@

To deal with pathologies such as the worst-case time for ML-style type
inference~\cite{ktu-caap-1990,m-popl-1990}, the typechecker
has internal limits that restrict the problems it will attempt to solve.
Hitting a limit triggers one of the following errors:
\code{CodeTooComplex},
\code{NormalizationTooComplex}, or
\code{UnificationTooComplex}.
These errors can appear in Script Analysis just like any other
error, but one design goal of Luau is that code rarely hits the limits.
A typical creator should never see these errors.

@@

  \mnocheck{}
  \includegraphics[width=\columnwidth]{img/error-count-nocheck-row--te-density-diff.pdf}

  \mnonstrict{}
  \includegraphics[width=\columnwidth]{img/error-count-nonstrict-row--te-density-diff.pdf}

  \mstrict{}
  \includegraphics[width=\columnwidth]{img/error-count-strict-row--te-density-diff.pdf}
  \caption{Changes in type error density ($\frac{\mbox{\#errors}}{\mbox{\#lines}}$) over time (seconds).}
  \label{f:error-density}
\end{figure}

@@

\Cref{f:error-density} plots changes to error density for every session in the dataset.
The x axis counts time in seconds from the start of the session.
Most sessions last a few dozen seconds~(\cref{t:session-size}), so the axis
ends at 150 seconds to show some but not all of the long tail.
The y axis shows changes to errors over time; a positive number means errors
increased since the last record, and a negative number represents a decrease.
While there are a few errors that go outside the plot bounds
(max \mnocheck{}: 101, max \mnonstrict{}: 3577, max \mstrict{}: 87), we focus
on $\pm{}50$ errors because most of the mass is in that region.
The points are color-coded by their session.
There are many colors througout the plot, which shows that the data comes
from many different sessions and is not dominated by a few high-traffic sessions.

@@

  \item
    After the 40-second mark, changes to type error density become much
    smaller.
    These long-running sessions are not introducing errors.
    %% TODO check #mod switch in long sessions

  \item
    \mnonstrict{} has the biggest fluctuations in errors, whereas \mstrict{}
    has the smallest.
    One explanation is that some \mnonstrict{} sessions ignore type analysis
    (in particular the one with over 3,000 errors!) whereas most \mstrict{}
    sessions pay close attention to it.

  \item
    When a color appears above the x axis, the same color usually appears below it as well.
    For example, \mnocheck{} has a cluster of purple points near the 20 second mark
    and $\pm 30$ density, which suggests that one session repeatedly added and removed errors.
    The next subsection conducts a thorough analysis.

@@

  %% with NON-moduleswitch data
% TE
%   nocheck & 961 & (\pct{38.50}) & 694 & [\pct{27.80}] & 841 & [\pct{33.69}] \\
% FS
%   nocheck & 156422 & (\pct{37.19}) & 153965 & [\pct{36.61}] & 110194 & [\pct{26.20}] \\
% TE mod
%   nocheck & 0 & (\pct{na}) & 0 & [\pct{na}] & 0 & [\pct{na}] \\
% FS mod
%   nocheck & 153913 & (\pct{36.96}) & 152735 & [\pct{36.68}] & 109807 & [\pct{26.37}] \\
% TE
%   nonstrict & 8623 & (\pct{37.67}) & 5263 & [\pct{22.99}] & 9004 & [\pct{39.34}] \\
% FS
%   nonstrict & 18927 & (\pct{37.41}) & 17791 & [\pct{35.17}] & 13872 & [\pct{27.42}] \\
% TE mod
%   nonstrict & 7185 & (\pct{36.61}) & 4768 & [\pct{24.30}] & 7672 & [\pct{39.09}] \\
% FS mod
%   nonstrict & 18573 & (\pct{37.09}) & 17604 & [\pct{35.15}] & 13905 & [\pct{27.76}] \\
% TE
%   strict & 442 & (\pct{38.27}) & 244 & [\pct{21.13}] & 469 & [\pct{40.61}] \\
% FS
%   strict & 296 & (\pct{31.66}) & 360 & [\pct{38.50}] & 279 & [\pct{29.84}] \\
% TE mod
%   strict & 251 & (\pct{35.35}) & 204 & [\pct{28.73}] & 255 & [\pct{35.92}] \\
% FS mod
%   strict & 272 & (\pct{32.04}) & 317 & [\pct{37.34}] & 260 & [\pct{30.62}] \\

>>>>

\newcommand{\etal}{et~al.}
\newcommand{\uid}{\textsc{id}}
\newcommand{\IDE}{\textsc{ide}}
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\FILL}{\textbf{FILL}}
\newcommand{\dotscale}[1]{\scalebox{0.72}{#1}}
\newcommand{\wideas}[2]{\makebox[\widthof{#2}][l]{#1}}
\newcommand{\twoline}[2]{\parbox[s]{1.4cm}{\flushleft#1\newline#2}}
\newcommand{\chkYes}{\dotscale{\CIRCLE}}
\newcommand{\chkMaybe}{\wideas{\dotscale{\Circle}}{\chkYes}}
\newcommand{\chkNo}{\wideas{}{\chkYes}}
\newcommand{\pct}[1]{\SI{#1}{\percent}}
\newcommand{\nanpct}{n/a}
\newcommand{\percentile}[1]{{#1}th~percentile}
\newcommand{\ptile}[1]{\textsc{p}{#1}}
\newcommand{\modefont}[1]{\texttt{#1}}
\newcommand{\mnocheck}{\modefont{nocheck}}
\newcommand{\mnonstrict}{\modefont{nonstrict}}
\newcommand{\mstrict}{\modefont{strict}}
\newcommand{\FS}{background}
\newcommand{\zerowidth}[1]{\makebox[0pt][l]{#1}}
\newcommand{\zeroheight}[1]{\raisebox{0pt}[0pt][0pt]{\(#1\)}}
\newcommand{\nK}[1]{#1\textsc{k}}
\newcommand{\stddev}[1]{\nK{#1}}
\newcommand{\gcell}[1]{\cellcolor{green!20}#1}
\newcommand{\ycell}[1]{\cellcolor{yellow!18}#1}
\newcommand{\ocell}[1]{\cellcolor{orange!29}#1}
\newcommand{\rcell}[1]{\cellcolor{red!30}$\!\!$#1$\!\!$}
\newcommand{\gbox}[1]{\colorbox{green!20}{#1}}
\newcommand{\ybox}[1]{\colorbox{yellow!18}{#1}}
\newcommand{\obox}[1]{\colorbox{orange!29}{#1}}
\newcommand{\rbox}[1]{\colorbox{red!30}{#1}}
\newcommand{\tefsfont}[1]{\textsc{#1}}
\newcommand{\tekey}{\tefsfont{te}}
\newcommand{\fskey}{\tefsfont{bg}}
\newcommand{\QALAN}{\FILL{} Alan:}
\newcommand{\panon}{pseudonymized}
\newcommand{\symbox}[1]{{\makebox[0.8em][c]{{#1}}}}
\newcommand{\addsym}{\symbox{\(\uparrow\)}}
\newcommand{\keepsym}{\symbox{\(=\)}}
\newcommand{\dropsym}{\symbox{\(\downarrow\)}}
% \newcommand{\dsymbox}[1]{\makebox[0.8em][c]{#1}}
% \newcommand{\dropsym}{\dsymbox{\raisebox{0ex}[0.9em]{\raisebox{0.5ex}{\vrule width 0.6em height 0.6pt}}}}

@@

Due to the broad community of creators, the goals of the Luau type
system are rather unique~\cite{bfj-hatra-2021}.
Whereas a traditional type system focuses on compilation and memory
safety, Luau takes a lenient approach by default and lets creators
gradually~\cite{st-sfp-2006,tf-popl-2008} migrate to rigorous checks
one module at a time.
In this way, for example,
students in code camps may begin in the \mnocheck{} mode that reports only
syntax errors; hobbyists can use \mnonstrict{} mode to identify
would-be runtime errors; and developers who need to ensure software
quality can rely on \mstrict{} mode.
Also, no matter the mode, a {\FS{}} type analysis runs to guide \IDE{} tools~(\cref{s:context}).
Luau's types aim to support untyped designs, in the spirit of migratory
typing~\cite{tfffgksst-snapl-2017}, so that creators can switch modes without
needing to restructure their code in major, potentially-breaking, ways.

@@

  \item
    An available dataset of over 1.5 million telemetry records
    and scripts to analyze them: \url{https://zenodo.org/doi/10.5281/zenodo.10275213}

@@

  \begin{tabular}{ll}
    Label & Interpretation \\\midrule
    \code{TypeMismatch} & Basic type error. \\
    \code{SyntaxError} & Basic parse error, e.g., \code{for if end}. \\
    \code{UnknownProperty} & Referenced an invalid field or method.  \\
    \code{OnlyTablesCanHaveMethods} & Tried to attach a method to a non-table. \\
    \code{CannotExtendTable} & Tried to extend a sealed table. \\
    \code{TypesAreUnrelated} & Failed cast, unify, or subtype. \\
    \code{CountMismatch} & Arity mismatch for a function. \\
    %% \code{CannotCallNonFunction} & Called a value that is not a function. \\
    \code{IncorrectGenericParamCount} & Arity mismatch for a generic type. \\
    \code{CodeTooComplex} & Type analysis failed~(\cref{s:code-too-complex}). \\
    %% \code{UnificationTooComplex} & Type analysis failed, unification solver hit a limit \\
    \code{GenericError}
    & Generic label for other non-type errors,\\
    & \hbox{}~~such as looping over an unordered table.
    %% attempting to extend a type that does not describe a class
    %% attempting to iterate over a table without a clear order

@@

  \item
    Telemetry \emph{must transmit a small amount of data} to avoid overloading the servers.
    This constrains both the number of records the system can transmit and the size of each record.

\end{enumerate}
See Cox~\cite{transparent-telemetry} for further discussion of
the privacy implications and tradeoffs in telemetry.

In addition, due to the architecture of Roblox Studio,
there is some information that is not available to the
typechecker:
\begin{itemize}
  \item
    Lifecycle events including save, run, quit, and publish.
    {(Publish events
    cannot be reported at any rate because they are public, and could be used
    to match a telemetry session to a creator.)}
  \item
    GUI state: in particular, whether the Script Analysis widget is visible.
\end{itemize}
Our type error telemetry has to meet all of the above constraints.
This means that it cannot send event-based telemetry on IDE lifecycle events
(due to Studio's architecture) and it cannot send full results from every type
analysis because that would overwhelm the data-collection server.

@@

\Cref{f:tdata} shows the structure of our telemetry records:
\begin{enumerate}
  \item
    \uid{}: a \panon{} identifier (15-digit random number) for the current session.
  \item
    Time: one timestamp from the client and one from the server.
    %% timestamp explanation \cref{s:data-cleaning}
  \item
    Mode: type analysis mode of the current file: \mnocheck{},
    \mnonstrict{}, or \mstrict{}.
  \item
    Reason: a flag that explains whether this record was sent due to a
    randomly-selected keystroke or a module switch.
  \item
    Size Info: number of lines in the codebase and number of lines in the edit range.
  \item
    Overall Counts: summary of type errors and \FS{} errors during the
    last two invocations of type analysis, and the number
    of times type analysis hit an internal limit on problem
    size~(see~\cref{s:code-too-complex} for details).
    For each invocation and each kind of
    analysis error, there are three summary counts:
    %% grand total = 12
    \begin{itemize}
      \item total number of errors,
      \item errors in the current module, and
      \item errors in the current edit range.
    \end{itemize}
  \item
    Edit Range Counts: list of specific errors that arose in the last two
    type analyses and overlap with the current edit range.
    Since there are 35 possible errors~(\cref{s:context}), there can be up to 70 counts
    in a telemetry record.
    Exactly how to interpret this data depends on which analysis it came from:
    \begin{itemize}
      \item
        Errors in the latest type analysis are currently visible to the
        creator.
        These may have been introduced by the changes in the edit range.
      \item
        Errors from the previous type analysis were visible before the creator
        made the latest edits.
        These may have motivated some of the changes in the edit range.
    \end{itemize}
\end{enumerate}

To track the edit range, we record a start and end position, which we
update appropriately on every edit. This can result in very large and
imprecise edit ranges, for example, if the user edits at the beginning and end
of the file.
We acknowledge this and other threats in~\cref{s:threats}.
The upshot of this strategy is that it reduces the size and complexity of
telemetry records because there is only one interval for errors to overlap with.

@@

\Cref{f:records-per-hour} provides a time-ordered distribution of the data.
The x-axis is labeled with a California time zone; the thin vertical lines
mark the start of a new day.
For each of the 24 hours in each day, there is one blue bar
indicating the number of records generated during that hour.
Each hour has a few dozen to over 3,000 records, typically about
1,000.
We conjecture the very low bars are due to outage on
the logging server and that the very tall bars are due to
user scripts that generated a huge number of keystrokes.
%% TODO investigate one tower, what's happening?? all one session?
Telemetry ended in early April; however, there are short bars in the mid-April
range from developers who waited a few days to update Roblox Studio.

The regular pattern of the peaks suggests that many creators
follow a regular schedule; these may be
professional developers.
The tall weekend (shaded) regions could be because
Roblox has a significant school-aged creator community.

@@

There is a huge amount of variation across sessions.
The largest ones have over 50,000 lines of code
while the smallest have zero lines of code.
%% TODO filter the zero lines? what happens in that data?? run a query
Unsurprisingly, these wide-ranging numbers come with large standard
deviations.
The median values are more reasonable, with roughly 3,000 lines of code
and 1,000 lines in edit ranges.

@@

\Cref{t:count-analysis-errors} lists the number of type analysis errors
and breaks them down by location in the codebase.
In total, there are over 590,000 type errors.
The number of \FS{} errors is much larger, at 72 million, because this analysis
uses rigorous checks and runs on every module no matter what type analysis
mode the module declares.
Furthermore, \FS{} checks run silently, so creators have no
awareness of or incentive to fix these errors.
Approximately half the analysis errors occurred in the current module,
and a small fraction of these overlapped with the current edit range~(\pct{5} for
type errors, \pct{1} for \FS{}).
The other errors point to different modules in the codebase.

The fraction of errors in the current module is relatively high because projects
contain many modules, all of which might contain latent type errors.
This high fraction is encouraging for two reasons.
First, it suggests that errors appear locally, as the result of edits to nearby
code, rather than as the result of edits to code that lives in another
module.
Second, it suggests that creators fix errors before switching
to another module.

The low fraction of errors in the edit range sends a mixed message:
\begin{itemize}
  \item
    On one hand, it may be that edits successfully remove errors.
    This is true, however, only for type errors and not for \FS{} errors.
    There are 11,479 more type errors from the previous analysis
    that overlap with the current edit range~(\pct{27} difference),
    but 172,678 fewer previous \FS{} errors~(\pct{15} difference).
    Since creators can see the type errors but not the \FS{}
    errors before making edits, we conclude that the visibility
    makes a difference.
    \Cref{s:type-error-survival} explores edits and type errors in more detail.

@@

The data does not support this explanation.
On average, mode-upgrades resulted in 3 additional
type errors (stddev 7, median 0).
This data excludes module switch records and considers only keystroke events.
The worst-case increase was quite high at 57 type errors, but exceptional.
In another exception, upgrading modes removed three type errors,
possibly due to other edits being rolled in with the
mode change.

@@

Digging further into \FS{} errors, the median number of such errors by mode is
1 in \mnocheck{}, 3 in \mnonstrict{}, and 1 in \mstrict{}.
Records in \mnocheck{} mode evidently do not have significantly more \FS{}
errors.
Furthermore, type analysis does not seem to reduce \FS{}
errors.
This is surprising;~\cref{s:tefs-compass} explores further.

@@

The main focus of our telemetry design is to learn how type errors intersect
with edit ranges.
Indeed, most of the fields in a telemetry record are dedicated to this
topic~(\cref{f:tdata}), and are directed to RQ2 on type errors .

@@

The counts are imprecise because of telemetry granularity.
If a creator removes one error but then
introduces another on a different line, this can manifest as
a keep (\keepsym{}) rather than one add (\addsym{)} and one drop (\dropsym{}).

@@

  \item
    \code{SyntaxError} and the related errors \code{UnknownSymbol} and
    \code{UnknownProperty} are the most common.
    If the edit range errors are indicative of type errors at large,
    then most errors in the data are likely due to typos during edits.
    These errors also survive (\keepsym{}) the more than any other.

    Curiously, the number of syntax errors never decreases or stays the same.
    This may be because the easy fix is to press undo, and telemetry cannot
    see undo events.

    %% TODO awkward
    Our telemetry strategy likely inflates the number of typos because it cannot
    determine when a creator is mid-edit.
    For example, a creator who writes a method call
    \code{bucket.countFish()} letter-by-letter will generate an
    \code{UnknownSymbol} error for \code{buck} and
    \code{UnknownProperty} for \code{bucket.c}, which may end up getting sampled.
    \Cref{s:conclusion} suggests ways
    to avoid this in the future.

@@

\begin{itemize}
  \item
    Typo-level errors, in particular \code{UnknownSymbol} (for an unbound variable),
    are much more common in \mnonstrict{} than \mstrict{}.
    This may be because \mnonstrict{} reports fewer errors
    to dilute the proportion of unknown symbols.

  \item
    Despite being less strict, \mnonstrict{} reported four errors that
    never appeared in \mstrict{} records:
    \code{DuplicateTypeDefinition}, \code{FunctionDoesNotTakeSelf},
    \code{MissingUnionProperty}, and \code{OnlyTablesCanHaveMethods}.
    We attribute this to the much larger number of \mnonstrict{} records in the
    dataset.

@@

To deal with pathologies such as the worst-case time for ML-style type
inference~\cite{ktu-caap-1990,m-popl-1990}, the typechecker
has internal limits that restrict the problems it will attempt to solve.
Hitting a limit triggers one of the following errors along with a message
that the analysis failed to understand the code:
\code{CodeTooComplex},
\code{NormalizationTooComplex}, or
\code{UnificationTooComplex}.
These errors can appear in Script Analysis just like any other
error, but one design goal of Luau is that code rarely hits the limits.
A typical creator should never see these errors.

@@

  \mnocheck{}
  \includegraphics[width=\columnwidth]{img/error-count-nocheck-row--te-density-diff-no-switch.pdf}

  \mnonstrict{}
  \includegraphics[width=\columnwidth]{img/error-count-nonstrict-row--te-density-diff-no-switch.pdf}

  \mstrict{}
  \includegraphics[width=\columnwidth]{img/error-count-strict-row--te-density-diff-no-switch.pdf}
  \caption{Changes in type error density ($\frac{\mbox{\#errors}}{\mbox{\#lines}}$) over time (seconds).}
  \label{f:error-density}
\end{figure}

@@

\Cref{f:error-density} plots changes to error density for every session in the dataset.
The data excludes module-switch telemetry records to keep the focus on edits.
The x axis counts time in seconds from the start of the session.
Most sessions last a few dozen seconds~(\cref{t:session-size}), so the axis
ends at 150 seconds to show some but not all of the long tail.
The y axis shows changes to errors over time; a positive number means errors
increased since the last record, and a negative number represents a decrease.
While there are a few errors that go outside the plot bounds
(max \mnocheck{}: 51, max \mnonstrict{}: 3586, max \mstrict{}: 87), we focus
on $\pm{}50$ errors because most of the mass is in that region.
The points are color-coded by their session.
There are many colors througout the plot, which shows that the data comes
from many different sessions rather than from a few high-traffic sessions.

@@

  \item
    Only a small fraction of \mnocheck{} records have nonzero type error density.
    This is in line with~\cref{t:type-error-survival}.
    Module switches in \mnocheck{} have nonzero density more often, but still
    rarely (\pct{3}).
    %% 34372 / 1337552 = 100 * 0.025664048949124968

  \item
    After the 40-second mark, changes to type error density become much
    smaller.
    These long-running sessions are not introducing errors.

  \item
    \mnonstrict{} has a bigger fluctuations in errors than \mstrict{}.
    One explanation is that some \mnonstrict{} sessions ignore type analysis
    (in particular the one with over 3,000 errors!) whereas most \mstrict{}
    sessions pay close attention to it.

@@

  %% with NON-moduleswitch data
  % TE nocheck & 961 & (\pct{38.50}) & 694 & [\pct{27.80}] & 841 & [\pct{33.69}] \\
  % TE nonstrict & 8623 & (\pct{37.67}) & 5263 & [\pct{22.99}] & 9004 & [\pct{39.34}] \\
  % TE strict & 442 & (\pct{38.27}) & 244 & [\pct{21.13}] & 469 & [\pct{40.61}] \\

  % TE mod nocheck & 0 & (\pct{na}) & 0 & [\pct{na}] & 0 & [\pct{na}] \\
  % TE mod nonstrict & 7185 & (\pct{36.61}) & 4768 & [\pct{24.30}] & 7672 & [\pct{39.09}] \\
  % TE mod strict & 251 & (\pct{35.35}) & 204 & [\pct{28.73}] & 255 & [\pct{35.92}] \\

  % FS nocheck & 156422 & (\pct{37.19}) & 153965 & [\pct{36.61}] & 110194 & [\pct{26.20}] \\
  % FS nonstrict & 18927 & (\pct{37.41}) & 17791 & [\pct{35.17}] & 13872 & [\pct{27.42}] \\
  % FS strict & 296 & (\pct{31.66}) & 360 & [\pct{38.50}] & 279 & [\pct{29.84}] \\

  % FS mod nocheck & 153913 & (\pct{36.96}) & 152735 & [\pct{36.68}] & 109807 & [\pct{26.37}] \\
  % FS mod nonstrict & 18573 & (\pct{37.09}) & 17604 & [\pct{35.15}] & 13905 & [\pct{27.76}] \\
  % FS mod strict & 272 & (\pct{32.04}) & 317 & [\pct{37.34}] & 260 & [\pct{30.62}] \\

  \begin{tabular}{l@{~~}lr@{}rr@{}rr@{}r}
    & & \zerowidth{\rbox{\addsym{}}} & & \zerowidth{\ybox{\keepsym{}}} & & \zerowidth{\gbox{\dropsym{}}} \\\midrule
    type error  & \mnocheck{}   &      0 & [\nanpct{}  ] &      0 & [\nanpct{}]   &      0 & [\nanpct{}] \\
    (in module) & \mnonstrict{} &   7185 & [\pct{36.61}] &   4786 & [\pct{24.30}] &   7672 & [\pct{39.09}] \\
                & \mstrict{}    &    251 & [\pct{35.35}] &    204 & [\pct{28.73}] &    255 & [\pct{35.92}] \\
    \\[-2ex]
    \FS         & \mnocheck{}   & 153913 & [\pct{36.96}] & 152735 & [\pct{36.68}] & 109807 & [\pct{26.37}] \\
    error       & \mnonstrict{} &  18573 & [\pct{37.09}] &  17604 & [\pct{35.15}] &  13905 & [\pct{27.76}] \\
    (in module) & \mstrict{}    &    272 & [\pct{32.04}] &    317 & [\pct{37.34}] &    260 & [\pct{30.62}] \\
  \end{tabular}
  \caption{How often do edits increase, maintain, or decrease the number of errors?}
  \label{f:error-changes}
\end{figure}
